---
title: "Analysis of the FCCSS"
output:
  html_document:
    df_print: paged
---

```{r}
library(dplyr)
library(GGally)
library(corrplot)
library(Rtsne)
library(PerformanceAnalytics)
library(ggplot2)
library(FactoMineR)
library(gridExtra)
```

##### Loading data:

-   We remove NAs

-   Discard patient identifiers

-   Remove patients who haven't received radiotherapy (Mean_dose_heart \> 0)

-   Compute the log survival time

-   Summarize the chemotherapy groups in one single variable that indicates if the patient underwent chemotherapy

```{r}
chemo <- c("ALKYL", "ANTHRA", "VINCA", "ANTIM", "ANTIB", "CHIMIO_GROUPE_AUTRE", "Chemotherapy")

grade3 <- read.csv("fccss.csv", header = TRUE, sep = ",", dec = ".") %>% 
  na.omit() %>%
  select(-"ctr", -"numcent") %>%
  mutate(log_survival_time = log(survival_time_years)) %>%
  filter(Pathologie_cardiaque_3_new == 1, Mean_dose_heart > 0 , Mean_dose_heart <20) %>%
  select(-Pathologie_cardiaque_3_new) %>%
  rowwise() %>%
  mutate(Chemotherapy  = as.integer(any(c_across(ALKYL:CHIMIO_GROUPE_AUTRE))))
  
grade3_chemo <-  grade3 %>%
  filter(Chemotherapy == 1) %>%
  select(-Chemotherapy)

grade3_no_chemo <- grade3 %>%
  filter(Chemotherapy == 0) %>%
  select(-Chemotherapy)

```

## Analysis:

#### Scatter plots

-   **Chemo + Radio :**

    ```{r}
    ggplot(data = subset(grade3_chemo, ALKYL==1), aes(y = survival_time_years , x = Mean_dose_heart)) + 
      geom_point() + 
      geom_smooth(method = "loess", formula = 'y~x')+
      ggtitle("Chemotherapy and radiotherapy ALKYL")


    ggplot(data = subset(grade3_chemo, ANTHRA==1), aes(y = survival_time_years , x = Mean_dose_heart)) + 
      geom_point() + 
      geom_smooth(method = "loess", formula = 'y~x')+
      ggtitle("Chemotherapy and radiotherapy ANTHRA")


    ggplot(data = subset(grade3_chemo, VINCA==1), aes(y = survival_time_years , x = Mean_dose_heart)) + 
      geom_point() + 
      geom_smooth(method = "loess", formula = 'y~x')+
      ggtitle("Chemotherapy and radiotherapy VINCA")


    ggplot(data = subset(grade3_chemo, ANTIM==1), aes(y = survival_time_years , x = Mean_dose_heart)) + 
      geom_point() + 
      geom_smooth(method = "loess", formula = 'y~x')+
      ggtitle("Chemotherapy and radiotherapy ANTIM")


    ggplot(data = subset(grade3_chemo, ANTIB==1), aes(y = survival_time_years , x = Mean_dose_heart)) + 
      geom_point() + 
      geom_smooth(method = "loess", formula = 'y~x')+
      ggtitle("Chemotherapy and radiotherapy ANTIB")


    ggplot(data = subset(grade3_chemo, CHIMIO_GROUPE_AUTRE==1), aes(y = survival_time_years , x = Mean_dose_heart)) + 
      geom_point() + 
      geom_smooth(method = "loess", formula = 'y~x')+
      ggtitle("Chemotherapy and radiotherapy AUTRE")
    ```

-   **Radio :**

    ```{r}
    ggplot(data = grade3_no_chemo, aes(y = survival_time_years, x = Mean_dose_heart)) + 
      geom_point() + 
      geom_smooth(method = "loess", formula = 'y~x')+
      ggtitle("Radiotherapy")
    ```

#### T-SNE

```{r}
#tsne_grade3 <- Rtsne(grade3, dims = 2, perplexity = 30, theta = 0.5)
tsne_grade3_chemo <- Rtsne(grade3_chemo, dims = 2, perplexity = 30, theta = 0.5)
tsne_grade3_no_chemo <- Rtsne(grade3_no_chemo, dims = 2, perplexity =11, theta = 0.5)
#plot(tsne_grade3$Y, main = "Grade3 chemo or radio")
plot(tsne_grade3_chemo$Y, main = "Chemotherapy and radiotherapy")
plot(tsne_grade3_no_chemo$Y, main = "Radiotherapy")
```

#### PCA:

**Radio + Chemo**:

```{r}
#pca_grade3 <- prcomp(grade3, scale. = TRUE)
pca_grade3_chemo <- PCA(grade3_chemo, graph = FALSE)
#biplot(pca_grade3)
plot(pca_grade3_chemo, choix = "var")
plot(pca_grade3_chemo, choix = "ind")
```

**Chemo:**

```{r}
pca_grade3_no_chemo <- PCA(grade3_no_chemo, graph = FALSE)
plot(pca_grade3_no_chemo, choix = "var")
plot(pca_grade3_no_chemo, choix = "ind")
```

#### Correlation matrix

**Radio:**

```{r message=FALSE, warning=FALSE}
ggpairs(grade3_no_chemo[c("survival_time_years", "age_at_diagnosis", "Mean_dose_heart", "log_survival_time")])
```

**Radio + Chemo**

```{r message=FALSE}
ggpairs(grade3_chemo[c("survival_time_years", "age_at_diagnosis", "Mean_dose_heart", "log_survival_time")])
```

```{r}
source("MixIdentifier.R")
source("rnornix.R")
Mix <- MixIdentifier(grade3$log_survival_time, 100000)
sim_log_sv <- rnormix(dim(grade3)[1], p = mean(Mix$mixture_wight), sigma_1 = mean(Mix$sigma1),mu_1 = mean(Mix$mu1), sigma_2 = mean(Mix$sigma2),mu_2 = mean(Mix$mu2))$samples
```

```{r}
ggplot() + 
  geom_histogram(aes(grade3$log_survival_time), bins = 100, alpha=.5, fill = "blue")+
  geom_histogram(aes(sim_log_sv), bins = 100, alpha=.5, fill= "green")+
  xlab("")
```

```{r}
ggplot() + 
  geom_histogram(aes(grade3$survival_time_years), bins = 50, alpha=.5, fill = "blue")+
  geom_histogram(aes(exp(sim_log_sv)), bins = 50, alpha=.5, fill= "green")+
  xlim(0, 100)
```
