---
title: "log-linear prediction model"
output: html_notebook
---

#### The model:

We consider a random variable Y that follows a Gaussian mixture

$$
Y \sim p\mathcal{N}(\mu_1, \sigma_1) + (1-p)\mathcal{N}(\mu_2, \sigma_2)
$$

Given the predictors $X = (X_1, \ldots, X_k)$ we express the overall mean of the model such that,

$$
\mathbb{E}[Y] = \mu = \log\left(\sum\beta_iX_i\right)
$$

and express the other parameters using Kamary $\textit{et al.}$ reparametrization

$$
\begin{align*} \mu_1=\mu-\sigma~\phi~.\left(\frac{\sqrt{1-p}}{\sqrt{p}}\right); &\quad \quad  \mu_2=\mu+\sigma~\phi~.\left(\frac{\sqrt{p}}{\sqrt{1-p}}\right)\\  \sigma_1=\sigma\left(\frac{\eta_1}{\sqrt{p}}\right); \quad &\quad \sigma_2=\sigma\left(\frac{\eta_2}{\sqrt{1-p}}\right)\\ \text{where}~\eta_1^2&+\eta_2^2+\phi^2=1\end{align*}
$$

### Data simulation:

To simulate data from the previous model, we start by simulating $n$ observations from $k$ covariates $X = (X_1, \ldots, X_k)$ using a multivariate normal distribution (Any other distribution can be used) and add the intercept $X_0$.

```{r}
n <- 100
k <- 4
set.seed(1999)
library(mvtnorm)
X <- matrix(data = NA, nrow = n, ncol = k)
X <- rmvnorm(n, mean = c(1,4,6,2))
X <- cbind(rep(1,n), X)
```

The coefficients $\beta_0, \dots, \beta_k$ for the linear predictor are chosen as follows,

$$
\mu  = \log(3X_0 + 10X_1 + 4X_2 + 9X_3 + 12X_4)
$$

```{r}
beta <- matrix(c(3,10,4,9,12), nrow = 1)
```

We now compute the overall mean of the model using the the log-linear relation

```{r}
mu <- log(beta %*% t(X))
```

For the choice of $\eta_1,\eta_2 \text{ and } \phi$ we use a randomly generated sample form the Dirichlet distribution

```{r}
library(DirichletReg)
set.seed(1999)
coeff <- sample(c(-1, 1), size = 1, replace = TRUE, prob = c(.5, .5))
etaphi <- c(1,1,coeff) * rdirichlet(1, c(.5, .5, .5))
eta1 <- etaphi[1]
eta2 <- etaphi[2]
phi <- etaphi[3]
```

The last parameters to choose are the mixture weight $p$ and the standard deviation $\sigma$

```{r}
p <- 0.2
sigma <- matrix(data = rep(1, n), nrow = 1)
```

Using all the previous parameters we compute the parameters of the mixture

```{r}
mu1 <- mu - sigma * phi * (sqrt(1-p)/sqrt(p))
mu2 <- mu + sigma * phi * (sqrt(p)/sqrt(1-p))
sigma1 <- sigma * (eta1/ sqrt(p))
sigma2 <- sigma * (eta2/sqrt(1-p))
```

we now simulate the response variable,

```{r}
source("rnornix.R")
set.seed(1999)
Y <- rnormix(n, p, mu1, sigma1, mu2, sigma2)$samples
```